---
title: "SEDA Maps"
author: "Amy Luo"
date: "11/20/2021"
output:
  html_document:
    fig_width: 10
    fig_height: 10
    theme:
      bootswatch: 'cosmo'
    df_print: paged
  pdf_document: default
---
### Congressional District, Political Leaning, and School Funding

**School Expenditure data** are available from: https://exhibits.stanford.edu/data/catalog/db586ns4974

**Political Ideology data** are available from: https://americanideologyproject.com/

- For further information on the measurement methodology for constituent policy preferences, please see this paper by [Tausanovitch and Warshaw (2013)](https://americanideologyproject.com/JOP_Tausanovich_Warshaw_2013.pdf) 

Variables of interest include:\

- *leaid* : NCES Local Education Agency (District) Code\
- *year* : spring of school year\
- *grade* : grade level\
- *cdcode*, *cd_fips* : congressional district code\
- *ppexp_tot* : total per pupil expenditures; tot exp/enrolled\
- *ppexp_inst* : instruction expenditures; instr exp/enrolled\
- *mrp_mean* : mean policy preferences of citizens in every congressional district estimates using multilevel regression with poststratification

#### Loading the data
```{r message=FALSE, warning=FALSE}
# Packages
library(haven)
library(dplyr)
library(USAboundaries)
library(dplyr)
library(sf)
library(leaflet)
library(dplyr)
library(sp)
library(ggplot2)
library(stringr)
library(dplyr)
library(kableExtra)
```

```{r message=FALSE, warning=FALSE}
# Loading SEDA dataset
full_data_yr <- read_dta("district by year by grade covariates from acs and ccd master_v1_1.dta")

# New data frame of variables of interest
dat <- full_data_yr %>% # filtering to 2013
  filter(year ==2013) 

dat <- dat[!duplicated(dat$leaid),] %>% # keeping only one observation per school district (removing grade level layer)
  select('leaid','cdcode','ppexp_tot','ppexp_inst') %>% 
  group_by(cdcode) %>% 
  mutate(ppexp = mean(ppexp_tot)) %>% # average per pupil expenditure in a congressional district
  mutate(piexp = mean(ppexp_inst)) %>% # average per instructor expenditure in a congressional district
  ungroup()

seda <- dat[!duplicated(dat$cdcode),] %>% # keeping only one observation per congressional district (removing school district layer)
  select(-'leaid',-'ppexp_tot',-'ppexp_inst') # dropping school district ID
```

#### Mapping Constituent Ideology by Congressional District
```{r message=FALSE, warning=FALSE}
# Loading USAboundaries from GitHub
devtools::install_github("ropensci/USAboundaries")
devtools::install_github("ropensci/USAboundariesData")

# Using USAboundaries package, load the spatial data 2010-2020 Congressional Districts
spatial <- us_congressional()


# Loading congressional ideology data for the 113th Congress from The American Ideology Project, https://americanideologyproject.com/
ideology <- read.csv('cd_113_TW_ideology_estimates.csv')


# Joining political ideology data with the spatial data
## String processing spatial dataset to remove leading zeros in "geoid"
spatial <- spatial %>% # Removing Alaska, Hawaii, and Puerto Rico
  mutate(cd_fips = str_remove(spatial$geoid, "^0+")) %>% 
  filter(state_abbr != 'AK') %>%
  filter(state_abbr != 'HI') %>%
  filter(state_abbr != 'PR')

## Selecting columns of interest from ideology dataset
ideology <- ideology %>%
  select(c('cd_fips', 'mrp_mean', 'pres_2012')) %>%
  mutate(cd_fips = as.character(cd_fips))

map <- left_join(spatial, ideology, by = 'cd_fips')

# Mapping
ideology_map <- ggplot() +
  geom_sf(data = map, 
          aes(geometry=geometry, fill=mrp_mean), alpha=2) +
  scale_fill_distiller(palette='RdBu', 
                       limits = c(-1.2,1.2), 
                       direction = -1, 
                       breaks=c(-1,0,1),
                       labels=c('Left','Center','Right'),
                       name='Political Leaning') +
  theme(panel.grid.major = element_blank(), 
        panel.background = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.subtitle = element_text(size=9, color='grey45')) +
  labs(title='Contituent Political Ideology by Congressional District', 
       subtitle = 'Source: The American Ideology Project (Tausanovitch & Warshaw, 2015)')

ideology_map
```

#### Background: Party Platform Positions on Education 
```{r message=FALSE, warning=FALSE}
tabd <- read.csv('Amy_Table.csv')
tabd %>%
  kbl(caption="Comparison between Dem and GOP platforms for K-12 education",
      format= "html",
      col.names = c("Topic","Democrat","Republican"),
                  align="l",) %>%
  column_spec(2, background = 'aliceblue') %>%
  column_spec(3, background = 'mistyrose') %>%
  row_spec(0, bold =T, background='lightgrey', color='black') %>%
  kable_paper(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
              full_width = F,
              html_font = "calibri",
              fixed_thead = T,
              position = 'center') %>%
  column_spec(1,bold=T, border_right =T) %>%
  footnote(general='https://www.ascd.org/blogs/what-the-democrats-and-republicans-stand-for-on-education-excerpts-from-party-platforms',
           general_title="Source: ") %>%
  scroll_box(width = "100%", height = "650px")
```


#### Distribution of Education Funding
```{r message=FALSE, warning=FALSE}
# Joining education dataset with map dataset
seda <- seda %>% 
  mutate(cd_fips=as.character(cdcode)) %>%
  select(-'cdcode')

map <- left_join(map, seda, by = 'cd_fips') 

# Per pupil Congressional District expenditure
map$z <- scale(map$ppexp)

# Per instructor Congressional District Expenditure
map$z_i <- scale(map$piexp)

# Creating indicator variable for histogram layers
hist <- map %>% 
  mutate(left = ifelse(mrp_mean>0, 'Right', 'Left')) # indicator variable for right (1) and left (0)

hist <- hist[!(is.na(hist$left)), ] # removing 1 na value

# Distribution of Student expenditures by Political Ideology
hist_p <- ggplot(hist, aes(x=z,color=left, fill=left)) +
  geom_histogram(position='identity', alpha = 0.3) +
  scale_color_manual(values=c('steelblue','salmon')) +
  scale_fill_manual(values=c('steelblue','salmon')) +
  labs(title='Distribution of Per Pupil Expediture',
       x='RMS value',
       y='Count',
       fill="Political Leaning") +
  guides(color=F) +
  theme_light(); hist_p

# Distribution of Instructor expenditures by Political Ideology
hist_i <- ggplot(hist, aes(x=z_i,color=left, fill=left)) +
  geom_histogram(position='identity', alpha = 0.3) +
  scale_color_manual(values=c('steelblue','salmon')) +
  scale_fill_manual(values=c('steelblue','salmon')) +
  labs(title='Distribution of Per Instructor Expediture',
       x='RMS value',
       y='Count',
       fill="Political Leaning") +
  guides(color=F) +
  theme_light(); hist_i
```

#### Mapping Average District Spending by Congressional District
```{r message=FALSE, warning=FALSE}
# Subsetting for map layers 
right <- map %>%
  subset(mrp_mean >0)
left <- map %>%
  subset(mrp_mean <0)

# Mapping per pupil Congressional District expenditure
palette <- colorBin("Greens", domain = c(-1.5,7), bins = 9)

leaflet(map) %>%
  addProviderTiles(providers$CartoDB.Positron, group='Political') %>%
  addTiles(group='Physical') %>%
  addPolygons(data=right,
               opacity=0.5,
               weight=2,
               color='salmon',
               fillColor = ~palette(z),
              fillOpacity = 1,
               highlight=highlightOptions(weight=3,
                                        color='aliceblue',
                                        bringToFront=T),
              group="Politically Right") %>%
  addPolygons(data=left,
              opacity=0.5,
              weight=2,
              color='steelblue',
              fillColor = ~palette(z),
              fillOpacity = 1,
              highlight=highlightOptions(weight=3,
                                      color='aliceblue',
                                      bringToFront=T),
               group='Politically Left') %>%
  addLayersControl(baseGroups=c('Political', 'Physical'),
    overlayGroups=c('Politically Right', 'Politically Left'),
    options=layersControlOptions(collapsed = F))%>%
  addLegend('bottomright',
            pal=palette,
            values=~z,
            opacity=0.5,
            title='Per Pupil Expenditure, scaled') 
```


```{r message=FALSE, warning=FALSE}
# Mapping per instructor Congressional District Expenditure
palette_i <- colorBin("Greens", domain = c(-1.3,7), bins = 9)

leaflet(map) %>%
  addProviderTiles(providers$CartoDB.Positron, group='Political') %>%
  addTiles(group='Physical') %>%
  addPolygons(data=right,
              opacity=0.5,
              weight=2,
              color='salmon',
              fillColor = ~palette_i(z_i),
              fillOpacity = 1,
              highlight=highlightOptions(weight=3,
                                        color='aliceblue',
                                        bringToFront=T),
              group="Politically Right") %>%
  addPolygons(data=left,
              opacity=0.5,
              weight=2,
              color='steelblue',
              fillColor = ~palette_i(z_i),
              fillOpacity = 1,
              highlight=highlightOptions(weight=3,
                                      color='aliceblue',
                                      bringToFront=T),
               group='Politically Left') %>%
  addLayersControl(baseGroups=c('Political', 'Physical'),
    overlayGroups=c('Politically Right', 'Politically Left'),
    options=layersControlOptions(collapsed = F))%>%
  addLegend('bottomright',
            pal=palette_i,
            values=~z_i,
            opacity=0.5,
            title='Per Instructor Expenditure, scaled') 
```




## Extra Work to Delete
```{r eval=FALSE, include=FALSE}
ideology_map <- ggplot() +
  geom_sf(data = map, 
          aes(geometry=geometry, fill=mrp), alpha=0.9) +
  scale_fill_steps2(limits = c(-1.2,1.2)) +
  theme(panel.grid.major = element_blank(), 
        panel.background = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) ; ideology_map
```

```{r eval=FALSE, include=FALSE}
## Subsetting data into right and left 
left <- map %>%
  subset(mrp_mean<0)

right <- map %>%
  subset(mrp_mean>0)

leaflet(map) %>%
  addProviderTiles(providers$CartoDB.Positron, group='Political') %>%
  addTiles(group='Physical') %>%
  addPolylines(data=right,
               opacity=0.5,
               weight=2,
               color='salmon',
               highlight=highlightOptions(weight=3,
                                        color='aliceblue',
                                        bringToFront=T),
              group="Right") %>%
  addPolylines(data=left,
               opacity=0.5,
               weight=2,
               color='steelblue',
               highlight=highlightOptions(weight=3,
                                      color='aliceblue',
                                      bringToFront=T),
               group='Left') %>%
  addLayersControl(
    baseGroups=c('Political', 'Physical'),
    overlayGroups=c('Right', 'Left'),
    options=layersControlOptions(collapsed = F)) %>%
  addLegend('bottomright',
            colors=c('steelblue','salmon'),
            labels=c('Left','Right'),
            title='Congressional District') 
```

```{r eval=FALSE, include=FALSE}
test <- read.csv('shd_2012_TW_ideology_estimates.csv') # state house district
test1 <- read.csv('county_TW_ideology_estimates.csv') # presidential elections by county
test2 <- read.csv('states_parties_estimates.csv') # state parties estimates
test3 <- read.csv('cd_113_TW_ideology_estimates.csv') # 113th congressional district level
```

```{r eval=FALSE, include=FALSE}
plot(st_geometry(right), col='tomato1')
plot(st_geometry(left), col='steelblue2')
```

```{r eval=FALSE, include=FALSE}
## Using leaflet package to make map
leaflet(md_115) %>% 
    addTiles() %>% 
    addPolygons(color = "black",
                opacity = 1,
                weight = 1,
                fillColor = ~palette(margin_percent),
                fillOpacity = 1,
                label = ~paste0("District: ", cd116fp))

map <- leaflet(political_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = political_map$geometry,
              color = 'black',
              opacity = 1,
              weight = 1,
              fillColor = ~palette(mrp_mean),
              fillOpacity = 1) ; map

## example from 
georgia_map %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = manual_fill), color = "black")  +
  scale_fill_manual(name = "Cases per 100,000", values = pal) +
  coord_fixed(1.3) +
  theme(panel.grid.major = element_blank(), 
        panel.background = element_blank(),
        axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  facet_grid(. ~ date)
```

```{r eval=FALSE, include=FALSE}
ggmap() + 
  geom_sf(data=map,aes(fill=`mrp_mean`),inherit.aes=FALSE,alpha=0.9) +
  scale_fill_gradient(low = "blue", high = "red", limits=c(20,80)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```


```{r eval=FALSE, include=FALSE}
md_votes <- read.csv(system.file("extdata", "md-115.csv",
                                   package = "USAboundaries"),
                       colClasses = c("character", "integer", "character", "integer",
                                      "character", "integer", "integer", "integer",
                                      "integer", "integer", "integer",  "numeric",
                                      "numeric", "numeric", "numeric", "numeric"))
md_districts <- us_congressional(states = "MD")


md_115 <- md_districts %>% 
  left_join(md_votes, by = c("cd116fp" = "district")) 

md_115$margin_percent <- md_115$percentage_republican - 0.5

palette <- colorBin("RdBu", domain = c(-0.3, 0.3), bins = 7, reverse = TRUE)

leaflet(md_115) %>% 
    addTiles() %>% 
    addPolygons(color = "black",
                opacity = 1,
                weight = 1,
                fillColor = ~palette(margin_percent),
                fillOpacity = 1,
                label = ~paste0("District: ", cd116fp))
```

```{r eval=FALSE, include=FALSE}
# Setting congressional district boundaries
get_congress_map <- function(cong=112) {
  tmp_file <- tempfile()
  tmp_dir <- tempdir()
  zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts112.zip",cong)
  download.file(zp, tmp_file)
  unzip(zipfile = tmp_file, exdir = tmp_dir)
  fpath <- paste(tmp_dir, sprintf("districtShapes/districts112.shp",cong), sep = "/")
  st_read(fpath)
}

# Loading maps from the 112th Congress (2011-2013)
cd112 <- get_congress_map(112)


# Selecting specific districts
cd112_nj <- cd112 %>% 
            filter(STATENAME=="New Jersey") %>%
            mutate(DISTRICT = as.character(DISTRICT)) %>%
            select(DISTRICT)

trump_nj <- tibble(DISTRICT=as.character(1:12),
                   `Trump Vote`=c(36.1, 50.6, 51.4, 55.8, 
                                    48.8, 40.6, 47.5, 21.5,
                                    33.1, 12.8, 48.8, 31.8))
cd112_nj <- cd112_nj %>% left_join(trump_nj, by="DISTRICT")
cd112_nj


dm <- get_map(location="New Jersey",zoom=8)


ggmap() + 
  geom_sf(data=cd112_nj,aes(fill=`Trump Vote`),inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient(low = "blue", high = "red", limits=c(20,80)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
```

```{r eval=FALSE, include=FALSE}
# Creating a composite variable for school district 'need' 
dat$need <- dat$stutch_all + (dat$speced/dat$totenrl) + (dat$ell/dat$totenrl)

## Weighted avg. by total student enrolled by congressional district


# Creating a composite variable for school district 'support'

## Weight avg. by total student enrolled by congressional district


```

```{r eval=FALSE, include=FALSE}
# Pull out USA map data frame
us_map <- map_data("world2", "USA")
us_map %>% ggplot(aes(x = long, y = lat, group = group)) +
                  geom_polygon(fill = "blue4", color = "white") + 
                  coord_fixed(1.3)

# Without Alaska and Hawaii
usa <- map_data("usa")
usa %>% ggplot(aes(x = long, y = lat, group = group)) +
        geom_polygon(color = "purple", fill = NA)

# Showing individual states
us_states <- map_data("state")

us_states %>% ggplot(aes(x = long, y = lat, fill = region, group = group)) + 
              geom_polygon(color = "white") + 
              coord_fixed(1.3) 

us_states %>% ggplot(aes(x = long, y = lat, fill = region, group = group)) + 
              geom_polygon(color = "white") + 
              coord_fixed(1.3) +
              guides(fill = FALSE) # do this to leave off the color legend

# Showing counties
AllCounty <- map_data("county")
AllCounty %>% ggplot(aes(x = long, y = lat, group = group)) +
              geom_polygon(color = "red", fill = NA)

AllCounty %>% ggplot(aes(x = long, y = lat, group = group)) +
              geom_polygon(color = "red", fill = NA, size = .1 )
```



```{r eval=FALSE, include=FALSE}
# Loading USAboundaries from GitHub
devtools::install_github("ropensci/USAboundaries")
devtools::install_github("ropensci/USAboundariesData")

library(USAboundaries) 
congress <- us_congressional()
plot(st_geometry(congress))
title("Congressional district boundaries in California")
```