---
title: "Untitled"
author: "Amy Luo"
date: "11/9/2021"
output: pdf_document
---
## Congressional District, Political Leaning, and School District Investments/Resources

These data are available online from the https://exhibits.stanford.edu/data/catalog/db586ns4974
The data selected are from the "districts covariates.dta" file listed. 

Variables of interest include:
- leaid: NCES Local Education Agency (District) Code
- year: spring of school year 
- grade: grade level
- cdcode: congressional district code
- totenrl: number of students in grade
- nsch: number of schools in the district 
- speced: number of special ED students in district
- ell: number of english language learners in district
- corsup: number of instructional coordinators and supervisors
- elmgui: number of elementary guidance counselors
- stutch_all: pupil teacher ratio -- average all student's school
- ppexp_tot: total per pupil expenditures; tot exp/enrolled
- ppexp_inst: instruction expenditures; instr exp/enrolled

### Loading Packages 
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(usmap)
library(ggplot2)
library(readr)
library(lubridate)
library(maps)
library(dplyr)
library(dslabs)
library(stringr)
```


```{r}
getwd()
library(haven)
full_data <- read_dta("/Users/ames/Desktop/Fall '21/school/BST 260/Final Project/district covariates.dta") # does not include year
full_data_yr <- read_dta("/Users/ames/Desktop/Fall '21/school/BST 260/Final Project/district by year by grade covariates from acs and ccd master_v1_1.dta")
#dplyr::glimpse(dat)

# New dataframe of variables of interest
dat <- full_data_yr[, c('leaid','year','grade','cdcode', 'totenrl', 'nsch','speced','ell','corsup','elmgui','stutch_all','ppexp_tot','ppexp_inst')]
```


### Pulling out a map data frame
```{r}
# Pull out USA map data frame
us_map <- map_data("world2", "USA")
us_map %>% ggplot(aes(x = long, y = lat, group = group)) +
                  geom_polygon(fill = "blue4", color = "white") + 
                  coord_fixed(1.3)

# Without Alaska and Hawaii
usa <- map_data("usa")
usa %>% ggplot(aes(x = long, y = lat, group = group)) +
        geom_polygon(color = "purple", fill = NA)

# Showing individual states
us_states <- map_data("state")

us_states %>% ggplot(aes(x = long, y = lat, fill = region, group = group)) + 
              geom_polygon(color = "white") + 
              coord_fixed(1.3) 

us_states %>% ggplot(aes(x = long, y = lat, fill = region, group = group)) + 
              geom_polygon(color = "white") + 
              coord_fixed(1.3) +
              guides(fill = FALSE) # do this to leave off the color legend

# Showing counties
AllCounty <- map_data("county")
AllCounty %>% ggplot(aes(x = long, y = lat, group = group)) +
              geom_polygon(color = "red", fill = NA)

AllCounty %>% ggplot(aes(x = long, y = lat, group = group)) +
              geom_polygon(color = "red", fill = NA, size = .1 )
```
